---
title: "alcohol-related-admissions"
author: "retegre"
date: "06/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

We just need the tidyverse package.

```{r load-packages}
library(tidyverse)
```

## Read in data
Our data is saved in csv files, so we will use the read_csv function.

```{r read-data}
# alcohol-related hospital admissions
data <- read_csv("scotpho_data_extract (1).csv") 

# intermediate data zone info
intermediate_zone_codes <- read_csv("iz2011_codes_and_labels_21042020.csv") 

```


## Prepare data
The data in these open datasets is already clean, so we don't need to worry about missing or wrong values. It is also tidy. 
We will only select the data we want: data from intermediate zones, and a subset of the variables from each dataset.

```{r}
glimpse(data)
```


```{r}
iz_data <- data %>%
  filter(area_type == "Intermediate zone") %>%
  select(area_name,area_code,year,measure)
glimpse(iz_data)
```


```{r}
glimpse(intermediate_zone_codes)
```

```{r}
iz_info <- intermediate_zone_codes %>%
  select(IntZone,HBName)

```
```{r}
glimpse(iz_info)
```

## Join the datasets
Let's join the two datasets, so we know which health board each data zone belongs to.
We will also remove "NHS" from the health board name, and rename a couple of the columns.

```{r join-datasets}
admission_data <- left_join(iz_data,iz_info, by = c("area_code" = "IntZone"))
```


```{r join-datasets}
glimpse(admission_data)
```


```{r join-datasets}
admission_data <- admission_data %>%
  mutate(HBName = gsub("NHS ", "", HBName)) %>%
  rename(health_board = HBName,
         alcohol_admissions = measure)

glimpse(admission_data)
```

## Save dataframe
To save the dataframe to a CSV file, we use the write_csv function.

```{r save-data}
write_csv(admission_data,"alcohol_related_admissions.csv")
```

##Exercise 1 
Plot distributions for each health board in 2019
```{r}
all_board_data_distribution <- admission_data %>%                          
  filter(year == 2019)      # only 2019  
#histogram
all_board_data_distribution %>%                         
  ggplot(aes(x = alcohol_admissions)) + 
  geom_histogram(bins = 15) +
  facet_wrap(~health_board)
```
# Q-Q plot
```{r}
all_board_data_distribution %>%                          
  ggplot(aes(sample = alcohol_admissions)) +          
  geom_qq() + 
  geom_qq_line(colour = "blue") + 
  facet_wrap(~health_board)
```

# boxplot
```{r}
all_board_data_distribution %>%                          
  ggplot(aes(x = health_board,             
             y = alcohol_admissions)) +    
  geom_boxplot() + 
  scale_x_discrete(guide = guide_axis(n.dodge=3))+
  geom_jitter(size = 0.3, alpha = 0.1) + # add data points
  theme(legend.position = "bottom")
  # remove legend

```





